{\LARGE Basis Choice for Scalar-on-Function Regression \\ with Applications to Near-Infrared Spectroscopy}

### 2022-02-08 Paper/main.tex
Small cut from bspline basis

	\paragraph{Monomial Basis}
	A very simple idea to approximate functions in $\mathbb{L}^2[0,1]$ is to take inspiration from the well known Taylor expansion and to use monomials of increasing degree as a basis. For entire functions $f(t)$, e.g. polynomials, the exponential function or trigonometric functions, we can express the function as a potentially infinite, converging sum of weighted monomials.
	\begin{equation}\label{Taylor_expansion}
		f(t) = \sum_{i = 1}^{\infty}a_i t^i \quad \text{where} \quad a_i = \frac{f^{(i)}(0)}{i!}
	\end{equation}

@@

	\paragraph{Fourier Basis}
	In the same way the monomial basis is connected to the Taylor series, the Fourier basis corresponds to the Fourier series. The Fourier series can be used to decompose a periodic function into a weighted sum of trigonometric functions. Equation \ref{Fourier_Series} shows an example for a periodic function with a period of 1.
	\begin{equation}\label{Fourier_Series}
		s(x) = \frac{A_0}{2} + \sum_{i = 1}^{\infty} A_n \cos(2\pi i x - \phi_n) = \frac{a_0}{2} + \sum_{i = 1}^{\infty}\left[a_i \cos(2\pi i x) + b_i \sin(2\pi i x)\right]
	\end{equation}

	In its classical form, the series is represented in the so called amplitude-phase form. This, however, is impractical for the estimation procedures shown in the later parts of this paper due to the phase shift parameter. Therefore, rewriting the series in the sine-cosine form as shown above is necessary.

	The Fourier basis for $\mathbb{L}^2[0,1]$ is thus given by the following sequence of functions defined on $[0,1]$ directly corresponding to the terms of the sine-cosine form of the Fourier series.
	\begin{equation}
		\phi_{i}^{F}(x) = 
		\begin{cases}
			1 & \text{if} \quad i = 1\\
			\sqrt{2} \cos(\pi i x) & \text{if} \quad i \quad \text{is even} \\
			\sqrt{2} \sin(\pi (i-1)x) & \text{otherwise}
		\end{cases}
	\end{equation}

@@

	\paragraph{B-spline Basis} Following chapter 3.5 from \cite{ramsay_functional_2005}, splines are defined by first dividing an interval of interest $[\tau_0, \tau_L]$ into $L$ subintervals of non-negative length divided by a non-decreasing sequence of points $(\tau_l)_{l = 1,\dots, L-1}$ called knots. On each subinterval, a spline is a polynomial of chosen order $m = n+1$ where $n$ is its degree. If there are no multiplicities in the set of these knots, the polynomials on neighboring subintervals are matching derivatives up to order $m-2$ at the boundary knot $\tau_l$. A typical case of this that is often used in practive is an equidistant grid of knots. In some settings, however, it can be sensible to place multiple knots at the same value to replicate specific properties of the data structure allowing for a reduced number of matching derivatives at the corresponding knots. For the purposes of this paper, we will focus on the case of equidistant knots without multiplicity at inner knots.\\

@@

	This, however, does not really lead to a basis of $\mathbb{L}^2[0,1]$ as the closed span of this finite sequence of functions is not equal to $\mathbb{L}^2[0,1]$. To focus on specific approximation errors in the later parts of this paper, we will, however, assume that a B-spline basis representation of a function in $\mathbb{L}^2[0,1]$ will serve as a sufficient approximation for an appropriately chosen number of B-spline basis functions. 
	Even though, this approach is not theoretically exact, in practice, this is often a reasonable approach and yields satisfactory results in cases where the functional form of B-splines makes them an appropriate approximation tool. \\

	As the B-spline basis does not have infinitely many elements as the full Monomial or Fourier bases do, it is slightly misleading to speak of truncating the B-spline basis at a truncation parameter $L$ in later parts of this paper. For the sake of keeping the notation reasonably concise, we will still keep this notation. However, by convention, truncating a B-spline basis at truncation parameter $L$ shall mean using a B-spline basis consisting of $L$ functions from this point on.

>>>>

	\paragraph{Monomial Basis}
	A very simple idea to approximate functions in $\mathbb{L}^2[0,1]$ is to take inspiration from the well known Taylor expansion and to use monomials of increasing degree as a basis. For entire functions $f(t)$, e.g. polynomials, the exponential function or trigonometric functions, we can express the function as a potentially infinite, but converging sum of weighted monomials.
	\begin{equation}\label{Taylor_expansion}
		f(t) = \sum_{i = 1}^{\infty}a_i t^i \quad \text{where} \quad a_i = \frac{f^{(i)}(0)}{i!}
	\end{equation}

@@

	\paragraph{Fourier Basis}
	In the same way the monomial basis is connected to the Taylor series, the Fourier basis corresponds to the Fourier series, which can be used to decompose a periodic function into trigonometric functions. Equation \ref{Fourier_Series} shows an example for a function with a period of 1.
	\begin{equation}\label{Fourier_Series}
		s(x) = \frac{A_0}{2} + \sum_{i = 1}^{\infty} A_n \cos(2\pi i x - \phi_n) = \frac{a_0}{2} + \sum_{i = 1}^{\infty}\left[a_i \cos(2\pi i x) + b_i \sin(2\pi i x)\right]
	\end{equation}

	Typically, the series is represented in the so called amplitude-phase form. This, however, is impractical for the estimation procedures shown in the later parts of this paper due to the phase shift parameter. Therefore, rewriting the series in the sine-cosine form as shown above is necessary.	
	The Fourier basis for $\mathbb{L}^2[0,1]$ is thus given by the following sequence of functions defined on $[0,1]$ directly corresponding to the terms of the sine-cosine form of the Fourier series.
	\begin{equation}
		\phi_{i}^{F}(x) = 
		\begin{cases}
			1 & \text{if} \quad i = 1\\
			\sqrt{2} \cos(\pi i x) & \text{if} \quad i \quad \text{is even} \\
			\sqrt{2} \sin(\pi (i-1)x) & \text{otherwise}
		\end{cases}
	\end{equation}

@@

	\paragraph{B-spline Basis} Following chapter 3.5 from \cite{ramsay_functional_2005}, splines are defined by first dividing an interval of interest $[\tau_0, \tau_L]$ into $L$ subintervals of non-negative length divided by a non-decreasing sequence of points $(\tau_l)_{l = 1,\dots, L-1}$ called knots. On each subinterval, a spline is a polynomial of order $m = n+1$ where $n$ is its degree. If there are no multiplicities in the set of knots, the polynomials on neighboring subintervals share derivatives up to order $m-2$ at the boundary knot $\tau_l$. A typical case that is often used in practice is an equidistant grid of knots. In some settings, however, it can be sensible to place multiple knots at the same value to replicate specific properties of the data structure allowing for a reduced number of matching derivatives at the corresponding knots. For the purposes of this paper, we will focus on the case of equidistant knots without multiplicity at inner knots.\\

@@

	This, however, does not lead to a basis of $\mathbb{L}^2[0,1]$ as the closed span of this finite sequence of functions is not equal to $\mathbb{L}^2[0,1]$. To focus on specific approximation errors in the later parts of this paper, we will, however, assume that a B-spline basis representation of a function in $\mathbb{L}^2[0,1]$ will serve as a sufficient approximation for an appropriately chosen number of B-spline basis functions.

	As the B-spline basis does not have infinitely many elements as the full monomial or Fourier bases do, it is slightly misleading to speak of truncating the B-spline basis at a truncation parameter $L$ in later parts of this paper. For the sake of keeping the notation reasonably concise, we will still keep this notation. However, by convention, truncating a B-spline basis at truncation parameter $L$ shall mean using a B-spline basis consisting of $L$ functions from this point on.
